---
title: "Digital Inclusion in Ashfield District"
subtitle: "Maps to accompany the Findings and Recommendations report to Ashfield District Council (June 2020)"
date: "June 2020"
author: "Citizens Online"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: flatly
    css: co_flex.css
    favicon: favicon.png
    logo: icon_white.png
mainfont: Source Sans Pro
---

```{r setup, include=FALSE}

library(here)
library(dplyr)
library(flexdashboard)
library(leaflet)
library(readr)
library(stringr)
library(tmap)

full_data <- readRDS(here("rds_data/full_data.Rds"))
digitypes_sf <- readRDS(here("rds_data/digitypes_sf.Rds"))
community_sf <- readRDS(here("rds_data/community_sf.Rds"))
ofcom_data_sf <- readRDS(here("rds_data/ofcom_data_sf.Rds"))
ward_boundaries <- readRDS(here("rds_data/ward_boundaries.Rds"))
district_boundary <- readRDS(here("rds_data/district_boundary.Rds"))

star_icon <- tmap_icons(here("star_fill.png"), width = 24, height = 24)
tmap_mode("view")
tmap_options(basemaps = leaflet::providers$CartoDB.Positron,
             fontfamily = "Source Sans Pro")

district_boundary_line <- tm_shape(district_boundary, name = "District boundary") +
    tm_borders(col = "lawngreen", lwd = 4, alpha = 0.5)
  
ward_boundary_lines <- tm_shape(select(ward_boundaries, wd19nm), name = "Ward boundaries") +
    tm_borders(col = "gold1", lwd = 2, alpha = 0.8)

community_assets <- tm_shape(community_sf, name = "Community assets") +
    tm_markers(shape = star_icon, popup.vars = "type")
  
low_sfbb <- tm_shape(filter(ofcom_data_sf, no_sfbb_pc > 0), name = "Postcodes with premises unable to access SFBB") +
    tm_dots(size = "no_sfbb_pc", col = "red2", alpha = 0.5, perceptual = TRUE, style = "cont", scale = 0.4, popup.vars = FALSE)

digi_nos <- tm_shape(filter(digitypes_sf, digi_type == "DigiNo"), name = "DigiNo postcodes") +
    tm_symbols(col = "white",
               border.col = "grey20",
               border.lwd = 2,
               size = 0.1,
               shape = 21,
               alpha = 0.7,
               popup.vars = FALSE)

default_hide_groups <- c("Ward boundaries", "Community assets", "Postcodes with premises unable to access SFBB")

```


### **Age and digital exclusion**<br>Digital exclusion risk is particularly acute where older people are living alone and/or have lower income. Specific numbers can be seen in a pop-up when you click on an area.<br>*Additional data layers are available as overlays: these may be turned on or off using the layers tool at the top left.* {data-commentary-width=500}


```{r over65_tmap}


full_data_age <- full_data %>% 
  mutate(
    over65_pct = round(
      100*over65_population/(working_age_population+over65_population), 1
      )
    ) %>% 
  mutate_at(vars(lsoa11nm), ~ str_replace(., "^Ashfield ", "")) %>% 
  select(lsoa11nm, over65_population, wd19nm, over65_pct, one_person_household_aged_65_and_over, pc_aug_19)

age_map <- tm_shape(full_data_age, name = "Age by LSOA (%)") +
      tm_polygons("over65_pct",
              title = "% adult population aged 65+",
              alpha = 0.6,
              border.col = "grey50",
              border.alpha = 0.6,
              lwd = 1,
              palette = "Purples",
              style = "log10",
              popup.vars = c("Ward" = "wd19nm",
                             "% adult popn. aged 65+" = "over65_pct",
                             "Popn. aged 65+" = "over65_population",
                             "Older people living alone" = "one_person_household_aged_65_and_over",
                             "Pension Credit claims" = "pc_aug_19")
              ) +
  
  tm_shape(full_data_age, name = "Older people living alone") +
    tm_polygons("one_person_household_aged_65_and_over",
                title = "People aged 65+ living alone",
                alpha = 0.5,
                border.col = "grey50",
                border.alpha = 0.5,
                lwd = 1,
                palette = "Purples",
                style = "log10",
                popup.vars = c("Ward" = "wd19nm",
                             "% adult popn. aged 65+" = "over65_pct",
                             "Popn. aged 65+" = "over65_population",
                             "Older people living alone" = "one_person_household_aged_65_and_over",
                             "Pension Credit claims" = "pc_aug_19")
                ) +
  
  tm_shape(dplyr::slice_max(full_data_age, order_by = pc_aug_19, n = 5), name = "Pension Credit claims") +
    tm_polygons(col = "white", alpha = 0, border.col = "deeppink", lwd = 2, border.alpha = 0.7, title = "Areas with highest Pension Credit claims") +
  
  district_boundary_line +
  ward_boundary_lines +
  community_assets +
  low_sfbb

tmap_leaflet(age_map) %>% 
  leaflet::hideGroup(c("Population aged 65+", default_hide_groups))

```

***

* Some LSOAs in the north of the district have high proportions of older residents, particularly in **Skegby** ward and in **Stanton Hill & Teversal**.
* In the south, the village of **Underwood** and one area in **Hucknall North** also have relatively old populations.
When the map of community assets is overlaid, we can see that these areas are also not well served by community centres or halls.
* While age is generally associated with digital exclusion, particular groups of older people are even more likely to be excluded - those who live alone and those who are on lower incomes.
* When we examine the distribution of older people living alone, we find a different pattern, with these being more likely to live in the towns, especially in Sutton and in the centre of Hucknall, as well as in Kirkby (**Summit** and **Kingsway** wards).
* The highest numbers of Pension Credit claimants (older people on lower incomes) are found in the same areas.
(The 5 LSOAs bordered with pink are those with the highest numbers of Pension Credit claimants.)
Being more urban, these areas have community centres and potential support venues not too far away, but very few actually in the highest-risk LSOAs themselves.
* Data from [Ofcom][ofcom] showing postcodes where some or all premises cannot access Superfast broadband (SFBB) is available as an overlay. Larger red dots represent higher proportions of premises affected in each postcode. Postcodes with full SFBB availability are not shown.
* Areas with poorer broadband availability are found throughout the district but with particular clusters in **Larwood**, **Annesley & Kirkby Woodhouse** and **Stanton Hill & Teversal** wards.
* Sources:
  * [ONS Mid-year population estimates (2018)][ons_popn]
  * [ONS Census (2011)/Nomis - Household composition][ons_hholds]
  * [DWP Stat-Xplore][statx]
  * [Ofcom Connected Nations 2019][ofcom]
  * [ONS OpenGeography][ons_geog]
  * [ONS Postcode Directory][ons_pc]

[ons_popn]: https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/lowersuperoutputareamidyearpopulationestimates
[ons_hholds]: https://www.nomisweb.co.uk/census/2011/qs113ew
[statx]: https://stat-xplore.dwp.gov.uk/webapi/home
[ofcom]: https://www.ofcom.org.uk/research-and-data/multi-sector-research/infrastructure-research/connected-nations-2019/data-downloads
[ons_geog]: https://geoportal.statistics.gov.uk/
[ons_pc]: https://geoportal.statistics.gov.uk/search?collection=Dataset&sort=-modified&tags=all(PRD_ONSPD)


### **Disability and digital exclusion**<br>Disabled people are much less likely to be online than non-disabled people. The map shows claimant numbers for three disability-related benefits (DLA, ESA, PIP).<br>*Additional data layers are available as overlays: these may be turned on or off using the layers tool at the top left.* {data-commentary-width=500}


```{r disability_tmap}

full_data_dis <- full_data %>% 
  mutate(
    disab_pct = round(
      100*comb_disability/(working_age_population+over65_population), 1
      )
    ) %>% 
  mutate_at(vars(lsoa11nm), ~ str_replace(., "^Ashfield ", "")) %>% 
  select(lsoa11nm, wd19nm, comb_disability, disab_pct, HDDDec)

disab_map <- tm_shape(full_data_dis, name = "Disability prevalence (% popn) (approx.)") +
      tm_polygons("disab_pct",
              title = "% adult population receiving disability benefits",
              alpha = 0.7,
              border.col = "grey50",
              border.alpha = 0.7,
              lwd = 1,
              palette = "Purples",
              style = "cont",
              n = 4,
              popup.vars = c("Ward" = "wd19nm",
                             "% adults receiving disability benefits (approx.)" = "disab_pct",
                             "Adults receiving disability benefits (approx.)" = "comb_disability",
                             "HDD Decile" = "HDDDec")
              ) +

    tm_shape(filter(full_data_dis, HDDDec %in% 1), name = "LSOAs with highest Health & Disability Deprivation") +
    tm_polygons(col = "white", alpha = 0, border.col = "deeppink", lwd = 2, border.alpha = 0.4, title = "LSOAs in highest HDD decile") +
  
  district_boundary_line +
  ward_boundary_lines +
  community_assets +
  low_sfbb

tmap_leaflet(disab_map) %>% 
  leaflet::hideGroup(default_hide_groups)

```

***

* While there are significant numbers of disabled people in every LSOA, the highest numbers are found particularly in Sutton (in **Carsic** and **Leamington** wards) and to a lesser extent in Kirkby (**Summit** and **Abbey Hill** wards).
* When the community assets layer is overlaid, we can see that in a few areas such as **Abbey Hill** and in the southern part of **Leamington** ward, there are very few or no places where disabled residents might be able to access digital skills support within their neighbourhood.
* The data shown here incorporates claims data from the DWP ([StatXplore][statx]) for the disability-related benefits DLA, ESA and PIP.
The claimant numbers are summed together here, but these benefits are not mutually exclusive, so the figures must be taken as approximations.
Additionally, not everyone who is disabled in a way that limits their day-to-day activities may be claiming one of these benefits.
* The [2019 English Indices of Deprivation][eimd] includes an index of health and disability deprivation, where LSOAs are scored and ranked nationally.
Those LSOAs in Ashfield within the 10% most deprived on this index are highlighted, for comparison (n = 10).
* Among these, three LSOAs (in **Huthwaite & Brierley**, **Central & New Cross**, and **Skegby**) are found outside the three target wards identified by the Citizens Online analysis.
* These areas may therefore have increased digital inclusion needs, and venues such as the nearby Huthwaite, Sutton and Skegby libraries (respectively), and for example New Cross Community Church, may be able to host digital support sessions with specific support for disabled residents.
* Sources:
  * [DWP Stat-Xplore][statx]
  * [2019 English Indices of Deprivation][eimd]
  * [ONS OpenGeography][ons_geog]

[eimd]: https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019




### **Low income and digital exclusion**<br>This map shows the number of Housing Benefit and Universal Credit claimants for each LSOA, as a measure of low income.<br>*Additional data layers are available as overlays: these may be turned on or off using the layers tool at the top left.* {data-commentary-width=500}


```{r income_tmap}

full_data_inc <- full_data %>% 
  mutate(hbuc = uc_claimants+hb_claimants) %>% 
  mutate(
    incom_pct = round(
      100*hbuc/(working_age_population+over65_population), 1
      )
    ) %>% 
  mutate_at(vars(lsoa11nm), ~ str_replace(., "^Ashfield ", "")) %>% 
  select(lsoa11nm, wd19nm, hbuc, incom_pct, lsoa_hh_no_sfbb)

incom_map <- tm_shape(full_data_inc, name = "Income deprivation prevalence (% adult popn)") +
      tm_polygons("incom_pct",
              title = "% adult population receiving UC or HB",
              alpha = 0.7,
              border.col = "grey50",
              border.alpha = 0.7,
              lwd = 1,
              palette = "Purples",
              style = "cont",
              n = 4,
              popup.vars = c("Ward" = "wd19nm",
                             "% adult population receiving UC or HB" = "incom_pct",
                             "Adults receiving UC or HB" = "hbuc",
                             "Households unable to access Superfast broadband (approx.)" = "lsoa_hh_no_sfbb"
                             )
              ) +
  
  district_boundary_line +
  ward_boundary_lines +
  community_assets +
  low_sfbb

tmap_leaflet(incom_map) %>% 
  leaflet::hideGroup(default_hide_groups)

```

***

* The highest numbers of people on lower incomes, as represented by HB and UC claims, are found within parts of **Leamington**, **Summit**, **Central & New Cross**, **Abbey Hill** and **Carsic** wards.
* These are areas where residents may be less likely to be able to buy, repair or replace devices due to cost, and/or may not be able to afford a mobile data allowance or a decent broadband connection.

* Sources:
  * [DWP Stat-Xplore][statx]
  * [ONS OpenGeography][ons_geog]


### **Overall wards heatmap**<br>Data at LSOA level is combined to ward level via an algorithm, and then weighted according to the relative importance of different factors (e.g. age) in digital exclusion, to produce summary digital exclusion risk scores for each ward.<br>*Additional data layers are available as overlays: these may be turned on or off using the layers tool at the top left.* {data-commentary-width=500}


```{r ward_heatmap}

median_risk <- readr::read_csv(here("ward_heatmap_numbers.csv")) %>% 
  select(wd19nm = Ward, de_risk = `Weighted DE Risk`, digino = `DigiNo postcodes`) %>% 
  pull(de_risk) %>% 
  median()


heatmap <- readr::read_csv(here("ward_heatmap_numbers.csv")) %>% 
  select(wd19nm = Ward, de_risk = `Weighted DE Risk`, digino = `DigiNo postcodes`) %>% 
  left_join(ward_boundaries, .) %>% 
  select(wd19nm, de_risk, digino) %>% 
  tm_shape(name = "Ward heatmap") +
    tm_polygons(title = "Digital Exclusion risk",
                col = "de_risk",
                alpha = 0.8,
                border.col = "grey50",
                border.alpha = 0.1,
                lwd = 2,
                palette = "-RdYlGn",
                style = "cont",
                breaks = c(24, 36, 48, 60, 72, 84, 99),
                midpoint = median_risk,
                popup.vars = c("Ward" = "wd19nm",
                               "DE Risk" = "de_risk",
                               "DigiNo postcodes" = "digino")
                ) +
  
  district_boundary_line +
  ward_boundary_lines +
  community_assets +
  low_sfbb +
  digi_nos


tmap_leaflet(heatmap) %>% 
  leaflet::hideGroup(c("Community assets", "DigiNo postcodes", "Postcodes with premises unable to access SFBB"))


```

***

* This map illustrates the *relative*, overall digital exclusion risk at ward level in Ashfield. There are significant variations between areas within wards in terms of risk and in terms of the accessibility of local support options. **Leamington** ward is an example where digital exclusion risk factors are highly concentrated in some parts, with risk being considerably lower in other areas.
* Likewise, wards not ranked as high-risk overall will still have numbers of people who are digitally excluded. These residents may be spread around the area, or found in more concentrated pockets. An example of this is **Abbey Hill** ward, where some parts are at high risk even though the ward as a whole is at moderate risk.
* The community assets layer shows which areas are best served by local services and potential community support venues, and where there are gaps.
* The Ashfield District Council DigiTypes classification has been included here as an optional overlay. Those postcodes categorised as "**DigiNo**" - with residents least likely to access services digitally - are shown as white dots. The spread of these postcodes across the district emphasises the point that digitally excluded residents are found in all areas, not just in the urban "hotspots".
* **NB** LSOA boundaries and ward boundaries in Ashfield do not always match up. Each LSOA is allocated to a ward by the ONS using a "best fit" process, which means the data we obtain at LSOA level sometimes includes some streets that are not actually within the allocated ward.
* Sources:
  * Ashfield District Council DigiTypes research
  * [ONS Postcode Directory][ons_pc]
  * [ONS OpenGeography][ons_geog]

